Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-05-18 doc/mc-data.tex
More data
Update data
Include results from CD4

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks
     \item Reduces set of benchmarks to 180
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 124 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around $0{.}90\times$ number of dDNNF nodes
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 124 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}
     \end{itemize}
   \item Ran program {\sc d2p} to generate CPOG file for maximum of
     10,000 seconds.  Verifies file, generates unweighted count. If
     finds literal weights, also generates weighted count
     \begin{itemize}
     \item Completed for 108 instances
     \item Exceeded clause limit for 1 instance (Our programs represent a clause ID as a 32-bit signed value)
     \item Ran out of memory for 1 instance
     \item Timed out for 14 instances
     \item Was able to do one-sided verification for 9 of these.
     \end{itemize}
   \item For completed CPOG files, ran {\sc cpog-check} CPOG checker
     \begin{itemize}
     \item Tests all conditions for correct CPOG
     \item Most time spent performing RUP checks for clause addition and deletion
     \item Completed successfully for all 108 instances
     \item Time generally faster than {\sc d2p}. (min = 0.006, max = 1.38, harmonic mean = 0.107)
     \end{itemize}
\end{enumerate}

@@

@@

Ran \lean{} checker on the benchmarks yielding at most $10^7$ clauses, with time limit of 1000 seconds.
\begin{itemize}
\item 80 total
\item All ran to completion
\item All but three ran within factor of $8\times$.  Outliers were small ones
\item Harmonic mean = $5.94$.
\end{itemize}

@@

\begin{description}
\item[Lemmas:] The justification proof must show that the conjunction
  of the input clauses implies the extension variable associated with
  the root node.  It does so by showing that this property holds for
  every node and for every path from the root that node, relative to
  the partial assignment given by that path.  This could potentially
  require enumerating exponential numbers of paths to some nodes,
  effectively converting the graph into a tree.  Instead, the proof
  can contain a {\em lemma} for each node having indegree greater than
  one, showing that the root node is implied by a set of clauses.
  Some of these are the original input clauses, while some are {\em
    synthetic clauses} representing input clauses that have been
  reduced by eliminating false literals.  Given such a proof, the
  lemma can be instantiated for each parent node.  Proofs with lemmas
  incur some overhead---the synthetic clauses are declared as POG
  product nodes, and additional clauses must be added to show that,
  under a given context, the synthetic clauses are implied by some
  input or containing synthetic clause.  On the other hand, they
  guarantee that no node in the graph is the root of more than one
  justication proof.
\item[Literal grouping:] The generated POG representation will
  typically have a number of product nodes forming the conjunction of
  tens, or even hundreds, of literals.  These literals represent ones
  that we either implied by unit propagation or were added when
  assigning the complement to the corresponding variable led to a
  conflict.  Justifying this product node requires showing that each
  of these literals must hold under the given context.  When this
  cannot be done by unit propagation, the program must invoke a SAT
  solver to prove that the complemented value yields a conflict.
  Rather than doing so for each literal individually, the prover can
  create a new product node consisting of the literals that have not
  yet been validated and then prove that falsifying the extension
  variable associated with the product node yields a conflict.  
\end{description}

>>>>

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks
     \item Reduces set of benchmarks to 180
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 124 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around ranged between $0.35\times$
         and $1.0\times$ number of dDNNF nodes (harmonic mean =
         $0.81\times$).  Greater optimizations with smaller
         benchmarks.
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 124 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}
     \end{itemize}
   \item Ran program {\sc d2p} to generate CPOG file for maximum of
     10,000 seconds.  Verifies file, generates unweighted count. If
     finds literal weights, also generates weighted count
     \begin{itemize}
     \item Completed for 108 instances
     \item Exceeded clause limit for 1 instance (Our programs represent a clause ID as a 32-bit signed value)
     \item Ran out of memory for 1 instance
     \item Timed out for 14 instances
     \item Was able to do one-sided verification for 9 of these.
     \end{itemize}
   \item For completed CPOG files, ran {\sc cpog-check} CPOG checker
     \begin{itemize}
     \item Tests all conditions for correct CPOG
     \item Most time spent performing RUP checks for clause addition and deletion
     \item Completed successfully for all 108 instances
     \item Time generally faster than {\sc d2p}. (min = 0.006, max = 1.38, harmonic mean = 0.107)
     \end{itemize}
\end{enumerate}

@@

Ran \lean{} checker on the benchmarks yielding at most $10^7$ clauses, with time limit of 1000 seconds.
\begin{itemize}
\item 80 total
\item All ran to completion
\item All but three ran within factor of $8\times$.  Outliers were small ones
\item Harmonic mean = $5.43\times$.
\end{itemize}

@@

\section{Comparison with CD4 Toolchain}

\input{figure-runtime-cd4}

Ran D4 with certificate logging enabled (by specifying DRAT file).
Used 114 benchmarks for which regular D4 ran in under 1000 seconds.
Measure time for entire toolchains (compilation, proof generation,
checking).  Checking for CD4 includes running drat-trim.
\begin{itemize}
\item Completed 80 in under 1000 seconds for both
  \begin{itemize}
    \item 7 ran faster with POG framework, up to $2.8\times$; 73 ran slower, up to $119.8\times$.  Harmonic mean = $19.63\times$ slower.
  \end{itemize}
\item 2 completed in under 1000 seconds  with POGs but timed out with CD4.  Both had very few models and with long unsatisfiability proofs
\item 28 completed in under 1000 seconds with CD4 but either timed out or failed with POGs.
\item Remaining 4 exceeded 1000 seconds for both toolchains

\end{itemize}

@@

\end{description}
